Importance sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-04-13 draft_p2.tex
Overwrite intro of part IV.B. (GKTL) by intro of part IV.B.1. (was redundant)

For time-averaged observables, one can make use of a different algorithm: the \acf{gktl} algorithm~\cite{Giardina2006,Tailleur2007,Giardina2011}.
It was originally designed to compute large deviations rate functions of time-averaged dynamical observables, that describe the probability density of both typical of rare events in the limit of large
averaging time~\cite{ref_large_dev}.
By contrast with the \ac{ams} and \ac{tams}, the \ac{gktl} algorithm does not follow the strategy of splitting algorithms.
Instead, it implements \textit{importance sampling}, \textit{i.e.} sampling a modified distribution that is biased towards the rare events of interest.
The idea of importance sampling is very general and was used in many different contexts (see e.g.~\cite{Berg1992,Hartmann2002} and the general references~\cite{Bucklew2004,RubinoTuffin2009}). 
The \ac{gktl} algorithm performs importance sampling in the space of trajectories, which is relevant for out-of-equilibrium systems.

\subsubsection{The \ac{gktl} algorithm}
\label{sec:gktl_description}
Similarly to the \ac{tams} algorithm, the \ac{gktl} relies on the simulation of a ensemble of trajectories.
In this case however, they interact dynamically: at regular time intervals, some members of the ensemble are killed and some are cloned according to a weight which depends on the history of the replica.
The weights are chosen such that, after several iterations of the algorithm, generated trajectories are distributed according to a probability distribution that is biased in favour of trajectories with large values of the time-averaged observable of interest.
The \ac{gktl} algorithm belong to a family of algorithms known as <<Go with the winners>>~\cite{aldous_varizani,grassberger}.
Similar ideas have been applied in a wide range of fields over the past 50 years, under different names, depending of the specific application domain~\cite{delMoral_Feynman_Kac_book}.
The application of a Go with the winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006~\cite{Giardina2006}.
Over the past ten years, it has successfully been applied to investigate rare events in both stochastic~\cite{Giardina2006,Lecomte2007b,Garrahan2007,Hurtado2009b} and deterministic systems~\cite{Giardina2006,Tailleur2007}.

The \ac{gktl} algorithm consists in simulation an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{1 \leq n \leq N}$ starting from independant random initial conditions.
Similarly to section~\ref{sec:ams}, the total integration time of the trajectories is denoted $T_{a}$.
We consider an observable of interest $A(\mathbf{x}(t))$ and a cloning period $\tau$.
At times $t_{i}=i\tau$ (with $i=1,2,...,T_{a}/\tau$) we assign to each trajectory $n$ a weight $W_{n}^{i}$ defined as
\begin{equation}
W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}}{R_{i}}\,\,\,\mbox{with}\,\,\,R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}.\label{eq:Weight}.
\end{equation}
For each trajectory $\{\mathbf{x}_{n}\}_{0 \leq t_i}$, a random number of copies of the trajectory are generated.
This number is on average proportional to the weight $W_{n}^{i}$, such that the total number of trajectories produced at each event is equal to $N$.
For deterministic systems, a small perturbation is introduced for clones at times $t_i$, so that clones separate from their parent. 
The parameter  is chosen by the user in order to control the strength of the selection.
The free parameter $k$ sets the amplitude of the typical fluctuations within the biased ensemble of trajectory sampled by the algorithm.
A higher value of $k$ will give a higher weight to higher fluctuations, thus driving the sampling to more extreme events.

>>>>

For time-averaged observables, one can make use of a different algorithm: the \acf{gktl} algorithm~\cite{Giardina2006,Tailleur2007,Giardina2011}.
Similarly to the \ac{tams} algorithm, the \ac{gktl} relies on the simulation of a ensemble of trajectories.
In this case however, they interact dynamically: at regular time intervals, some members of the ensemble are killed and some are cloned according to a weight which depends on the history of the replica.
The weights are chosen such that, after several iterations of the algorithm, generated trajectories are distributed according to a probability distribution that is biased in favour of trajectories with large values of the time-averaged observable of interest.
The \ac{gktl} algorithm belong to a family of algorithms known as <<Go with the winners>>~\cite{aldous_varizani,grassberger}.
Similar ideas have been applied in a wide range of fields over the past 50 years, under different names, depending of the specific application domain~\cite{delMoral_Feynman_Kac_book}.
The application of a Go with the winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in 2006~\cite{Giardina2006}.
Over the past ten years, it has successfully been applied to investigate rare events in both stochastic~\cite{Giardina2006,Lecomte2007b,Garrahan2007,Hurtado2009b} and deterministic systems~\cite{Giardina2006,Tailleur2007}.
% It was originally designed to compute large deviations rate functions of time-averaged dynamical observables, that describe the probability density of both typical of rare events in the limit of large
% averaging time~\cite{ref_large_dev}.
% By contrast with the \ac{ams} and \ac{tams}, the \ac{gktl} algorithm does not follow the strategy of splitting algorithms.
% Instead, it implements \textit{importance sampling}, \textit{i.e.} sampling a modified distribution that is biased towards the rare events of interest.
% The idea of importance sampling is very general and was used in many different contexts (see e.g.~\cite{Berg1992,Hartmann2002} and the general references~\cite{Bucklew2004,RubinoTuffin2009}). 
% The \ac{gktl} algorithm performs importance sampling in the space of trajectories, which is relevant for out-of-equilibrium systems.

\subsubsection{The \ac{gktl} algorithm}
\label{sec:gktl_description}
The \ac{gktl} algorithm consists in simulation an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{1 \leq n \leq N}$ starting from independant random initial conditions.
Similarly to section~\ref{sec:ams}, the total integration time of the trajectories is denoted $T_{a}$.
We consider an observable of interest $A(\mathbf{x}(t))$ and a cloning period $\tau$.
At times $t_{i}=i\tau$ (with $i=1,2,...,T_{a}/\tau$) we assign to each trajectory $n$ a weight $W_{n}^{i}$ defined as
\begin{equation}
W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}}{R_{i}}\,\,\,\mbox{with}\,\,\,R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(X_{n}(t))dt}.\label{eq:Weight}.
\end{equation}
For each trajectory $\{\mathbf{x}_{n}\}_{0 \leq t_i}$, a random number of copies of the trajectory are generated.
This number is on average proportional to the weight $W_{n}^{i}$, such that the total number of trajectories produced at each event is equal to $N$.
For deterministic systems, a small perturbation is introduced for clones at times $t_i$, so that clones separate from their parent. 
The parameter  is chosen by the user in order to control the strength of the selection.
The free parameter $k$ sets the amplitude of the typical fluctuations within the biased ensemble of trajectory sampled by the algorithm.
A higher value of $k$ will give a higher weight to higher fluctuations, thus driving the sampling to more extreme events.
